% !TEX program = xelatex
% RAI Risk Taxonomy Technical Report 2.0 -- journal-style English edition.
\documentclass[11pt,a4paper]{article}
\usepackage[a4paper,top=22mm,bottom=24mm,left=22mm,right=22mm]{geometry}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage{booktabs,longtable,tabularx,array}
\usepackage{enumitem}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage[hidelinks]{hyperref}
\pgfplotsset{compat=1.18}

\setmainfont{Times New Roman}
\setsansfont{Arial}
\setmonofont{Menlo}
\setlength{\parindent}{1.4em}
\setlength{\parskip}{0pt}
\setlength{\emergencystretch}{2em}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\pct}[1]{#1\%}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\definecolor{natureblue}{HTML}{0072B2}
\definecolor{naturevermillion}{HTML}{D55E00}
\definecolor{naturegreen}{HTML}{009E73}
\definecolor{naturegrey}{HTML}{6B6B6B}
\definecolor{naturelight}{HTML}{D9D9D9}

\pgfplotsset{
  natureaxis/.style={
    axis x line*=bottom,
    axis y line*=left,
    axis line style={black, line width=0.6pt},
    tick align=outside,
    tick style={black, line width=0.5pt},
    tick label style={font=\sffamily\footnotesize, color=black},
    label style={font=\sffamily\small, color=black},
    title style={font=\sffamily\small, color=black},
    legend style={draw=none, fill=none, font=\sffamily\footnotesize},
    grid=none,
    clip=false
  }
}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\lhead{\footnotesize RAI Risk Taxonomy}
\rhead{\footnotesize Technical Report 2.0}
\cfoot{\footnotesize\thepage}
\hypersetup{
  pdftitle={RAI Risk Taxonomy Technical Report 2.0},
  pdfauthor={RAI Risk Taxonomy Project}
}

\title{\vspace{-1.3cm}\bfseries\Large RAI Risk Taxonomy 2.0:\\[3pt]
Top-down evidence selection and staged classification of 1,711 global AI risks}
\author{RAI Risk Taxonomy Project}
\date{21 July 2026}

\begin{document}
\maketitle

\begin{abstract}
We present a reproducible, top-down account of the evidence and classification
pipeline used to construct RAI Risk Taxonomy 2.0. The integrated evidence space
contains 3,906,767 AI records: 1,749,159 papers, 125,372 policy reports, and
2,032,236 patents. A curated AI-risk evidence view contains 82,971 records
(2.1\% of the AI evidence space). Canonicalization, source crosswalking, and
risk-card construction yields 1,726 candidate rows. A conservative exact-match
check merges one duplicate, yielding 1,725 candidate L4 cards. A taxonomy-level
check against the 24 authoritative Physical AI L3 labels excludes 14 exact
label leaks, yielding 1,711 L4 cards. A staged hierarchy-blind classification
procedure assigns all 1,711 cards (100.0\%) to 50 L3 nodes; 76 cards (4.4\%)
remain marked \code{HOLD}. We specify the
filter at each transition, the mathematical decision rules, the protected
Physical AI set, and the limits of interpreting record-level evidence as
card-level taxonomy coverage.
\end{abstract}

\noindent\textbf{Keywords:} AI risk taxonomy; evidence synthesis; open-set
classification; semantic mapping; Physical AI; responsible AI.

\section{Introduction}

A taxonomy can appear complete while concealing uncertainty in its source
corpus, unit of analysis, or placement rules. RAI Risk Taxonomy 2.0 therefore
separates four quantities that must not be conflated: the size of the integrated
AI evidence space, the size of the AI-risk evidence view, the number of
canonical L4 risk cards, and the number of L4 cards assigned to L3. The central
result is a top-down evidence-to-taxonomy chain of 3,906,767 records, 82,971
AI-risk records, 1,726 candidate L4 rows, 1,725 exact-deduplicated candidates,
1,711 level-validated L4 cards, and 1,711 L3-assigned cards.

The transitions do not all preserve the same statistical unit. The first two
stages count documents or patent records. The third and fourth stages count
canonical risk concepts. Consequently, the 82,971 evidence records are not a
one-to-one parent table from which every L4 card was mechanically generated.
They form an auditable evidence view supporting corpus coverage and risk
interpretation, whereas the L4 registry integrates risk taxonomies, benchmarks,
standards, and linked evidence through a source crosswalk.

\section{Top-down dataset construction}

\subsection{Stage A: integrated AI evidence space}

The top-level dataset contains 3,906,767 records from three source families:
1,749,159 papers, 125,372 policy reports, and 2,032,236 patents. Paper records
originate from the local global Web of Science AI collections. Policy records
combine the reviewed curated AI-policy master with the broader policy-materials
collection. Patent records originate from the stage-2 AI patent master.

Records are standardized to a common schema and retained only when a non-empty
title is available. Deduplication is performed within document type and
analytical category. The ordered key hierarchy is patent application ID for
patents, followed by DOI, source record ID, URL, and a normalized
title--year--institution fallback. Cross-category overlap is intentionally
retained because AI, AI infrastructure, Physical AI, and AI risk are analytical
views rather than mutually exclusive partitions.

\subsection{Stage B: integrated AI-risk evidence view}

The AI-risk evidence view contains 82,971 records: 50,375 papers, 938 policy
reports, and 31,658 patents. Its inclusion rule is source-governed rather than a
single keyword query. Papers combine curated academic risk references, the
clean and integrated AI-risk paper corpora, Physical AI safety benchmarks, and
verified paper records from the top-200 evidence set. Policy reports combine
curated risk-policy references with verified report records. Patents combine
curated risk-patent references with verified patent records. Record-type filters
are applied before the same title-validity and within-category deduplication
rules used at Stage A.

This view represents 2.1\% of the integrated AI evidence space. Because the
source collections and analytical categories can overlap, this percentage is a
descriptive ratio, not the estimated prevalence of risk research in all AI
scholarship.

\subsection{Stage C: canonical L4 registry}

The unit changes from evidence record to risk concept. Source risk IDs are
joined through the source crosswalk, semantically redundant cards are reviewed,
and one canonical row is retained per L4 failure mode or harm mechanism.
Permanent identifiers \code{RAI4-0001} through \code{RAI4-1726} are allocated
independently of the risk label. The initial result is 1,726 candidate L4 rows. Automatic
deduplication is restricted to pairs whose normalized label and mechanism-only
definition are both exactly equal; semantic-similarity thresholds are not used.
This merges \code{RAI4-1606} into \code{RAI4-1083}, retains both source references,
and yields 1,725 exact-deduplicated candidates. The retired ID remains in the
crosswalk. The registry includes the 182-card Physical AI taxonomy as a
protected mapping set.

The ratio of L4 cards to AI-risk evidence records is 2.1\%. It is reported only
as a scale-change diagnostic: one document may support multiple risks, and one
risk may be supported by multiple documents. It is not a document acceptance
rate.

\subsection{Stage D: taxonomy-level validation}

The deployed Physical AI site is the authoritative source for its 24 L3 names
and 182 L4 cards. Its live HTML was byte-identical to the local repository
snapshot used for this check. We compare the English component of each
authoritative Physical L3 name with every candidate L4 label after NFKC
normalization, case folding, ampersand normalization, and removal of
non-alphanumeric separators. A card is excluded only when these normalized
labels are exactly equal; semantic similarity and definition similarity are not
used. Fourteen L4 rows match category names and are retired as taxonomy-level
leakage, including \code{RAI4-0553} ``Accidental harm'', which maps to Physical
L3 \code{P3.1}. Their definitions and references remain in the audit crosswalk.
This yields 1,711 canonical L4 cards without changing any of the protected 182
Physical L4 mappings.

\subsection{Authoritative Physical L4 content synchronization}

Release v2.4 synchronizes the content of all 182 protected Physical cards from
the local repository underlying the deployed Physical AI Risk Taxonomy site.
The immutable RAI4 crosswalk and L3 paths are retained, while the bilingual
label and definition, severity, probability, 3H1R attributes, evidence
justifications, and 360 reference rows are replaced by the current Physical
source values. Of these reference rows, 359 contain a justification and one
source row retains an empty justification without imputation. Bilingual
terminal parenthetical text is split into Korean and
English fields to preserve the global card schema. Impact is recalculated as
$I=s\,p$ from the synchronized severity $s$ and probability $p$. Percentile is
not displayed for any card because the cards do not constitute a common
measurement population suitable for relative ranking. Existing percentile
fields remain provenance data only.

\subsection{Complete English--Korean localization}

Release v2.5 requires bilingual display at every taxonomy level. The supplied
L0--L3 terminology and Korean definitions are used as the controlled glossary.
The 182 Physical L4 cards retain their authoritative bilingual source text. For
the remaining 1,529 cards, the localization pipeline removes the repeated
provenance boilerplate from each English definition, selects the first sentence
that states the core failure mechanism, generates a concise Korean definition,
and normalizes risk-domain terminology. Context-sensitive corrections include
\textit{poisoning} as ``data or model contamination'' rather than substance
addiction, \textit{washing} as false compliance presentation, and
\textit{rubber stamping} as nominal approval. Every generated row retains the
English source-sentence hash, method, and \code{human\_review\_status=pending}.
Thus all 1,711 cards contain non-empty English and Korean labels and definitions,
while automated Korean drafts remain explicitly reviewable rather than being
represented as human-approved ground truth.

\subsection{Canonical L2 consolidation}

Release v2.6 normalizes L2 to three canonical categories:
\textit{Interaction Safety}, \textit{System Safety}, and \textit{Societal
Safety}. The former General labels \textit{Interaction} and \textit{System},
and the former Agentic label \textit{System}, are renamed accordingly. Six
L1-specific path nodes remain as implementation edges so that the strict
L0--L1--L2--L3 tree and all 50 existing L3 parent paths are preserved; each
edge references one of the three canonical L2 identifiers. Thus the reported
L2 category count is three, while no L3 path or L4 assignment changes.

\subsection{Stable L4 identifiers and L1 display semantics}

The public explorer keeps the stable \code{RAI4-\#\#\#\#} identifier independent
of taxonomy placement. L1 membership is communicated through an explicit text
badge, a consistent icon, and a visual color: a brain and blue for
General-purpose AI, a compass and green for Agentic AI, and a robot and red for
Physical AI. The icon--color pair is repeated in the global navigation,
filters, hierarchy panel, and L4 cards. The color is derived from the card's current L1
path at render time rather than encoded into the permanent identifier. This
allows later human-approved remapping without identifier churn and avoids using
color or icon as the sole carrier of categorical information.

\subsection{Stage E: L3 assignment policy}

The final policy view assigns all 1,711 canonical L4 cards to one of 50 L3 nodes
(100.0\%). After taxonomy-level validation, 76 cards (4.4\%) retain a \code{decision\_required}
marker, displayed as \code{HOLD}. This marker is metadata on an assigned card,
not an L1, L2, or L3 taxonomy node.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
  natureaxis,
  xbar,
  xmode=log,
  log origin=infty,
  log basis x=10,
  xmin=1000,
  xmax=20000000,
  width=0.62\linewidth,
  height=7.0cm,
  xlabel={Count (log scale)},
  symbolic y coords={L3-assigned L4,Level-validated L4,Exact-deduplicated candidates,AI-risk evidence,AI evidence},
  ytick={L3-assigned L4,Level-validated L4,Exact-deduplicated candidates,AI-risk evidence,AI evidence},
  yticklabels={L3-assigned L4,Level-validated L4,Exact-deduplicated candidates,AI-risk evidence,AI evidence},
  yticklabel style={font=\sffamily\scriptsize,text width=2.25cm,align=right},
  bar width=12pt,
  point meta=explicit symbolic,
  nodes near coords={\pgfplotspointmeta},
  nodes near coords style={font=\sffamily\footnotesize, anchor=west},
]
\addplot[fill=natureblue, draw=natureblue] coordinates {
  (1711,L3-assigned L4) [1,711]
  (1711,Level-validated L4) [1,711]
  (1725,Exact-deduplicated candidates) [1,725]
  (82971,AI-risk evidence) [82,971]
  (3906767,AI evidence) [3,906,767]
};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Top-down evidence and taxonomy funnel.} Counts move from
document/patent records to canonical risk concepts at the transition from
AI-risk evidence to L4. The logarithmic axis preserves visibility across four
orders of magnitude.}
\label{fig:funnel}
\end{figure}

\begin{table}[H]
\centering
\caption{Top-down stages, counts, and inclusion filters.}
\label{tab:funnel}
\small
\begin{tabularx}{\linewidth}{L{0.18\linewidth}r L{0.20\linewidth}Y}
\toprule
Stage & Count & Unit & Principal filter \\
\midrule
AI evidence & 3,906,767 & Document or patent record & AI source collections; valid title; standardized schema; within-type/category deduplication. \\
AI-risk evidence & 82,971 & Document or patent record & Curated/integrated risk sources; record-type inclusion; valid title; within-type/category deduplication. \\
Exact-deduplicated candidates & 1,725 & Risk card & Exact normalized label and core-definition match only; one retired ID mapped to its canonical card. \\
Canonical L4 & 1,711 & Risk card & Exact normalized-label exclusion against 24 authoritative Physical L3 names; 14 level leaks retired; all 182 Physical L4 cards preserved. \\
L3 assigned & 1,711 & Risk card & Protected Physical mapping or staged non-Physical placement; 76 decision-required cards remain assigned and marked \code{HOLD}. \\
\bottomrule
\end{tabularx}
\end{table}

\section{Mathematical classification procedure}

\subsection{Protected Physical AI mapping}

The 182 L4-to-L3 mappings inherited from the Physical AI taxonomy are treated
as the locked gold set. Later stages cannot change these paths. Before the
taxonomy-level exclusion, the remaining 1,543 candidates enter the
hierarchy-blind classification procedure. Legacy global
L0--L3 labels, source families, and source-ID prefixes remain reference metadata
and are excluded from prediction features.

\subsection{Stage 1: strict open-set gate}

The input for each non-Physical card is the L4 label, a mechanism-only
definition with hierarchy restatements removed, and the evidence title. Let
$s_{sem}$ be the top-1 semantic score, $s_{comp}$ the top-1 composite score,
$m_{sem}$ and $m_{comp}$ their respective top-1 margins, and $v_{anchor}$ the
number of anchor views selecting the same L3. A candidate is retained only when
all seven conditions hold:

\begin{enumerate}[leftmargin=2em]
  \item exactly one L3 is rule-eligible;
  \item the rule winner, semantic top-1, and composite top-1 are identical;
  \item $s_{sem}\geq0.55$ and $s_{comp}\geq0.54$;
  \item $m_{sem}\geq0.035$ and $m_{comp}\geq0.035$;
  \item $v_{anchor}\geq2$ across three anchor views;
  \item no gap sentinel, multi-mechanism flag, or Physical-boundary flag is present; and
  \item two independent expert agents approve the same exact L3.
\end{enumerate}

This gate yields 37 new expert-consensus proposals. Together with the 182
Physical locks, Stage 1 assigns 219 cards and withholds 1,507.

\subsection{Stage 2: suitability-ranked relaxation}

Stage 2 preserves all 219 protected assignments and computes
\begin{equation}
S_2=0.55q_{sem}+0.20q_{sm}+0.10q_{cm}+0.10q_{vote}+0.05q_{rule}-p_{gap},
\end{equation}
where
\begin{align}
q_{sem}&=\operatorname{clip}\left(\frac{s_{sem}-0.45}{0.25},0,1\right),\\
q_{sm}&=\operatorname{clip}\left(\frac{m_{sem}}{0.05},0,1\right),\\
q_{cm}&=\operatorname{clip}\left(\frac{m_{comp}}{0.05},0,1\right),\\
q_{vote}&=v_{anchor}/3,\\
q_{rule}&=\mathbf{1}[\text{at least one eligible L3}],\\
p_{gap}&=0.05\,\mathbf{1}[\text{gap sentinel present}].
\end{align}

$S_2$ is a deterministic ranking score rather than a calibrated probability.
Stage 2 first preserves 55 hard holds. It then orders the remaining holds by
ascending $S_2$, with permanent RAI4 ID as the tie breaker, and reserves the
lowest 118. Thus $55+118=173$ cards remain unresolved and 1,553 cards are
assigned (90.0\%).

\subsection{Stage 3: forced-path test and final policy}

Stage 3 preserves the 1,553 Stage 2 paths and assigns only the 173 unresolved
cards to the hierarchy-blind top-1 L3 already stored in the score artifact.
This produces 1,726 operational paths but does not validate taxonomy fit. Each
new path retains its original hold reason, score, review priority, and
\code{human\_approved=false} marker.

The final policy marks five difficult-case families for decision review:
23 Physical-scope reviews, 17 expert rejections, 7 expert disagreements,
6 multi-mechanism cases, and 2 low-absolute-fit cases. Their forced L3 path
remains populated and \code{decision\_required=true}. The
remaining 118 quota holds also retain their forced paths and review metadata.

\subsection{Independent expert card review}

Two reviewers independently inspected all 1,725 cards for agreement among the
L4 label, mechanism definition, evidence, and assigned L3. Reviewer A proposed
30 amendments and Reviewer B proposed 14. Automatic amendment required both
reviewers to propose the same label and L3 at high confidence. One card met this
rule: \code{RAI4-0888} changed from ``Privacy concerns'' to
``Anthropomorphic trust-induced privacy disclosure'', while retaining
\code{RAI3-G-INT-10}. Its original label remains in provenance metadata.
Forty-two cards had a proposal from only one reviewer or incompatible proposals;
their labels and paths were not changed. Thirteen were already marked \code{HOLD}
and 29 newly received the marker. The consensus amendment cleared the prior hold
on \code{RAI4-0888}, giving $55-1+29=83$ final \code{HOLD} cards. No protected
Physical AI card was changed.

The later taxonomy-level validation supersedes the proposed adjudication of
\code{RAI4-0553}: ``Accidental harm'' is itself the authoritative Physical L3
category \code{P3.1}, not an L4 failure mode. It is therefore excluded rather
than renamed or remapped. Thirteen additional exact Physical-L3 label leaks are
handled identically. Six of the excluded rows previously carried \code{HOLD},
reducing the displayed count from 82 to 76. The protected Physical set remains
exactly 182 cards.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
  natureaxis,
  xbar,
  xmin=0,
  xmax=125,
  width=0.74\linewidth,
  height=6.7cm,
  xlabel={Cards},
  symbolic y coords={Low absolute fit,Multiple mechanisms,Expert disagreement,Expert rejection,Physical scope review,Suitability reserve},
  ytick=data,
  bar width=10pt,
  nodes near coords,
  nodes near coords style={font=\sffamily\footnotesize, anchor=west},
  enlarge y limits=0.12,
]
\addplot[fill=naturevermillion, draw=naturevermillion] coordinates {
  (2,Low absolute fit) (6,Multiple mechanisms) (7,Expert disagreement)
  (17,Expert rejection) (23,Physical scope review)
};
\addplot[fill=natureblue, draw=natureblue] coordinates {(118,Suitability reserve)};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Disposition of the 173 Stage 2 holds.} The 55 difficult cases
(vermillion) are force-assigned and marked \code{HOLD}; 118 suitability-reserve
cards (blue) are also force-assigned.}
\label{fig:holds}
\end{figure}

\section{Results}

\subsection{Evidence composition}

Papers account for 44.8\% of the integrated AI evidence space, policy reports
for 3.2\%, and patents for 52.0\%. Within the AI-risk evidence view, the
corresponding shares are 60.7\%, 1.1\%, and 38.2\%. These shares use one decimal
because their integer component is at least one.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
  natureaxis,
  ybar,
  ymin=0,
  ymax=70,
  width=0.80\linewidth,
  height=6.3cm,
  ylabel={Share (\%)},
  symbolic x coords={Papers,Policy reports,Patents},
  xtick=data,
  bar width=12pt,
  legend columns=2,
  legend pos=north west,
]
\addplot[fill=naturegrey, draw=naturegrey] coordinates {(Papers,44.8) (Policy reports,3.2) (Patents,52.0)};
\addplot[fill=natureblue, draw=natureblue] coordinates {(Papers,60.7) (Policy reports,1.1) (Patents,38.2)};
\legend{AI evidence,AI-risk evidence}
\end{axis}
\end{tikzpicture}
\caption{\textbf{Source-family composition.} Percentages compare document-type
composition within the integrated AI evidence space and the AI-risk evidence
view.}
\label{fig:composition}
\end{figure}

\subsection{Final hierarchy coverage}

The final policy assigns all 1,711 canonical cards (100.0\%); 76 assigned cards (4.4\%)
carry the \code{HOLD} marker. Of all cards, 1,180 are General, 349 are Agentic
AI, and 182 are Physical AI. Their shares are 69.0\%, 20.4\%, and 10.6\%, respectively.
All 50 L3 nodes contain at least one card. Anthropomorphism is the largest L3
with 247 cards (14.4\%); the five largest L3 nodes jointly contain 41.9\%.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
  natureaxis,
  xbar,
  xmin=0,
  xmax=260,
  width=0.70\linewidth,
  height=9.2cm,
  xlabel={L4 cards},
  symbolic y coords={Weaponization,Non-Contestability,Conflict,Destabilising Dynamics,Excessive Authority,Network Effects,Privacy,Policy Exposure,Misinformation/Disinformation,Copyrights,Goal Misalignment,Anthropomorphism},
  ytick=data,
  bar width=8pt,
  nodes near coords,
  nodes near coords style={font=\sffamily\scriptsize, anchor=west},
]
\addplot[fill=natureblue, draw=natureblue] coordinates {
 (59,Weaponization) (55,Non-Contestability) (63,Conflict)
 (73,Destabilising Dynamics) (74,Excessive Authority) (79,Network Effects)
 (87,Privacy) (89,Policy Exposure) (110,Misinformation/Disinformation)
 (118,Copyrights) (153,Goal Misalignment) (247,Anthropomorphism)
};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Largest L3 nodes in the final policy view.} The figure shows
the twelve largest L3 nodes. Counts include the 76 assigned cards marked
\code{HOLD}.}
\label{fig:l3}
\end{figure}

\section{Rounding and reporting policy}

All counts are reported as integers. Percentages below 1 are rounded at the
third decimal place and displayed to two decimal places; percentages with an
integer component of at least 1 are rounded and displayed to one decimal place.
For example, the L4 registry represents 0.04\% of the integrated AI evidence
count, the AI-risk view represents 2.1\%, final L3 coverage is 100.0\%, and the
decision-required share is 4.4\%. Exact floating-point ratios remain available in the
machine-readable statistics artifact.

\section{Validation and limitations}

The 182 Physical mappings and 37 Stage 1 expert-consensus mappings remain
unchanged in later stages. Repeated Stage 2 and Stage 3 builds produce
byte-identical placement hashes. The Stage 2 validator reports 13 PASS, 0 FAIL,
and 3 WARN; the Stage 3 validator reports 14 PASS, 0 FAIL, and 4 WARN. The
release validator reports 71 PASS, 0 FAIL, and 2 WARN, and all 32 repository
tests pass at the time of this report revision.

Three limitations directly affect interpretation. First, cross-category
overlap is retained in the integrated evidence space, so category counts are
analytical views rather than mutually exclusive samples. Second, the transition
from 82,971 evidence records to 1,711 canonical L4 cards changes the unit of analysis and
is not a mechanical screening step. Third, 1,214 of the 1,334 new Stage 2
placements lack Stage 1 rule eligibility, 247 override a taxonomy-gap sentinel,
and the remaining 118 Stage 3 forced candidates are not human-approved ground
truth.

\section{Review priorities}

The 76 assigned cards marked \code{HOLD} should be reviewed first, beginning with
the 42 cards lacking two-reviewer consensus and the original hard-hold cases. The 118
quota candidates should then be reviewed in ascending $S_2$ order. Finally,
the 247 gap overrides and 1,214 rule-free placements should be audited through
L3-stratified samples. Human review should determine whether each failure is
best addressed by expanding an L3 definition, introducing a new L3, or allowing
primary and secondary labels.

\section*{Data availability}

The exact top-down counts, filters, placements, and public card bundle are
available in the repository:
\begin{itemize}[leftmargin=2em]
  \item \small\nolinkurl{reports/statistics/top_down_evidence_funnel.json}
  \item \small\nolinkurl{data/experiments/stage3-forced-policy-v2/placements.json}
  \item \small\nolinkurl{public/data/releases/v2.6.0/cards.json}
  \item \small\nolinkurl{reports/data_quality/l4_deduplication_v2.1/retired_to_canonical.json}
  \item \small\nolinkurl{reports/expert_review/v2.1/consensus_result.json}
  \item \small\nolinkurl{reports/data_quality/physical_l3_label_leakage_v2.3.json}
  \item \small\nolinkurl{reports/data_quality/physical_card_sync_v2.4.json}
  \item \small\nolinkurl{reports/localization/nonphysical_l4_ko_v2.5.json}
  \item \small\nolinkurl{reports/data_quality/l2_category_consolidation_v2.6.json}
\end{itemize}
This document is the English journal-style edition of \textit{RAI Risk
Taxonomy Technical Report 2.0}.

\end{document}
